---
title: "R Notebook"
output: html_notebook
---

```{r, setup}
knitr::opts_knit$set(root.dir = "1SNPperPeak_projectedSNPs/")
```


```{r, functions}
make_K<-function(rootname){
  main<-readGRM(rootname)
  main_K=matrix(0,max(main$grm$id1),max(main$grm$id2))
  main_K[upper.tri(main_K,diag=T)]=main$grm$grm
  main_K[lower.tri(main_K)]=t(main_K)[lower.tri(main_K)]
  return(main_K)
}

normalize_kinmat <- function(kinmat){
  #normalize kinship so that Kij \in [0,1]
  tmp=kinmat - min(kinmat)
  tmp=tmp/max(tmp)
  #fix eigenvalues to positive
  diag(tmp)=diag(tmp)-min(eigen(tmp)$values)
  return(tmp)
}


readGRM <- function(rootname)
{
  #rootname="MOA.SNPs.bed_CentIBS_K" #originally in Jeff's code, but I think it's overriding the input so everything is being read in the same way...
  bin.file.name <- paste(rootname, ".grm.bin", sep="")
  n.file.name <- paste(rootname, ".grm.N.bin", sep="")
  id.file.name <- paste(rootname, ".grm.id", sep="")
  
  cat("Reading IDs\n")
  id <- read.table(id.file.name, colClasses="character")
  n <- dim(id)[1]
  cat("Reading GRM\n")
  bin.file <- file(bin.file.name, "rb")
  grm <- readBin(bin.file, n=n*(n+1)/2, what=numeric(0), size=4)
  close(bin.file)
  cat("Reading N\n")
  n.file <- file(n.file.name, "rb")
  N <- readBin(n.file, n=n*(n+1)/2, what=numeric(0), size=4)
  close(n.file)
  
  cat("Creating data frame\n")
  l <- list()
  for(i in 1:n)
  {
    l[[i]] <- 1:i
  }
  col1 <- rep(1:n, 1:n)
  col2 <- unlist(l)
  grm <- data.frame(id1=col1, id2=col2, N=N, grm=grm)	
  
  ret <- list()
  ret$grm <- grm
  ret$id <- id
  return(ret)
}
```


```{r}
library("tidyverse")
```

```{r}
moa_K<-make_K("MOA.SNPs.bed_CentIBS_K") 
smoa_K<-svd(moa_K)

back_K<-make_K("BKGD.SNPs.bed_CentIBS_K") 
sback_K<-svd(back_K)

rest_K<-make_K("CentIBS_Rest_K") 
srest_K<-svd(rest_K)
```

```{r}
for(rnd in 1:10){
# Three K matrices plus residuals
cmoa_K = sqrt(smoa_K$d) * t(smoa_K$u) # chol(moa_K)
cback_K = sqrt(sback_K$d) * t(sback_K$u) 
crest_K = sqrt(srest_K$d) * t(srest_K$u) 
# create 4 vectors of random variables with each kinship matrix (or uncorrelated residuals)
U = matrix(0,nrow = nrow(cmoa_K),ncol = 4)
U[,1] = crossprod(cmoa_K,rnorm(nrow(cmoa_K)))
U[,2] = crossprod(cback_K,rnorm(nrow(cback_K)))
U[,3] = crossprod(crest_K,rnorm(nrow(crest_K)))
U[,4] = rnorm(nrow(cmoa_K))
# Create a grid of variance component proportions for the 4 variance components
#grid = seq(0,1,by = 0.2)
#h2s = expand.grid(h21 = grid,h22 = grid,h23 = grid)
#h2s$h24 = 1-rowSums(h2s)
#h2s = h2s[h2s$h24 > 0,]

# If we want very specific variance component proportions to test
h2s = tibble(h21 = c(rep(0.1,1),rep(0.5,1),rep(0.05,1),rep(0.9, 1)), 
             h22 = c(rep(0.5,1),rep(0.1,1),rep(0.05,1),rep(0.02, 1)), 
             h23 = c(rep(0.1,1),rep(0.1,1),rep(0.3,1),rep(0.02, 1))
             )
h2s$h24 = 1-rowSums(h2s)
# for each row of h2s, create a y by re-weighting the vectors of random variables
Y = U %*% t(sqrt(h2s))
apply(Y,2,sd)
# column i of Y is a trait with variance component proportions h2s[i,]
colnames(Y)<- c(paste("Set",1:4,"-",rnd,sep = ""))
if(rnd == 1){ #iteratively adds the new trait to a main dataframe
  Sim_Traits = as.data.frame(Y) } else{
    Sim_Traits = add_column(Sim_Traits, Y[,1], Y[,2],Y[,3],Y[,4])
  }
colnames(Sim_Traits)[(ncol(Sim_Traits)-3):ncol(Sim_Traits)] = colnames(Y) #changes the column names from "Y" to trait name
}
write_tsv(Sim_Traits, paste("Simulated_Phenotypes",".","Round2",".txt",sep = ""))
```
Write Y as the phenotype output file (if not doing the loop)
```{r}
colnames(Y)<- c(paste("Set",1:4,"-","1",sep = ""))
write_tsv(as.data.frame(Y), paste("Simulated_Phenotypes",".1",".txt",sep = ""))
```
#Trouble shooting why the simulations gave weird results for set 1 and set 3
Set 2 and Set 4 simulations seemed to be ok, but 1 and 3 were really off. 
So I'll make new phenotypes for set 1 and set 3 since I can't seem to find a labeling error. 
I'll make the phenotypes individually so there's no chance of a mis-labeling
***Later realized that 1 Kinship matrix was being read in for all 3. Error fixed in Round 2 simulations***
```{r}
for(rnd in 1:10){
# Three K matrices plus residuals
cmoa_K = sqrt(smoa_K$d) * t(smoa_K$u) # chol(moa_K)
cback_K = sqrt(sback_K$d) * t(sback_K$u) 
crest_K = sqrt(srest_K$d) * t(srest_K$u) 
# create 4 vectors of random variables with each kinship matrix (or uncorrelated residuals)
U = matrix(0,nrow = nrow(cmoa_K),ncol = 4)
U[,1] = crossprod(cmoa_K,rnorm(nrow(cmoa_K)))
U[,2] = crossprod(cback_K,rnorm(nrow(cback_K)))
U[,3] = crossprod(crest_K,rnorm(nrow(crest_K)))
U[,4] = rnorm(nrow(cmoa_K))
# Create a grid of variance component proportions for the 4 variance components
#grid = seq(0,1,by = 0.2)
#h2s = expand.grid(h21 = grid,h22 = grid,h23 = grid)
#h2s$h24 = 1-rowSums(h2s)
#h2s = h2s[h2s$h24 > 0,]

# If we want very specific variance component proportions to test
#set1
#h2s = tibble(h21 = c(rep(0.1,1),rep(0.5,1),rep(0.05,1),rep(0.9, 1)), 
#             h22 = c(rep(0.5,1),rep(0.1,1),rep(0.05,1),rep(0.02, 1)), 
#             h23 = c(rep(0.1,1),rep(0.1,1),rep(0.3,1),rep(0.02, 1))
#             )
h2s = tibble(h21 = c(rep(0.05,1)), 
             h22 = c(rep(0.05,1)), 
             h23 = c(rep(0.3,1))
             )
h2s$h24 = 1-rowSums(h2s)
# for each row of h2s, create a y by re-weighting the vectors of random variables
Y = U %*% t(sqrt(h2s))
apply(Y,2,sd)
# column i of Y is a trait with variance component proportions h2s[i,]
colnames(Y)<- c(paste("Set",3,"-",rnd,sep = "")) #makes the trait name
if(rnd == 1){ #iteratively adds the new trait to a main dataframe
  Set3_Traits = as.data.frame(Y) } else{
    Set3_Traits = add_column(Set3_Traits, Y)
  }
colnames(Set3_Traits)[rnd] = colnames(Y) #changes the column names from "Y" to trait name
}
#write_tsv(Set1_Traits, paste("Simulated_Phenotypes",".","Set1only",".txt",sep = ""))
write_tsv(Set3_Traits, paste("Simulated_Phenotypes",".","Set3only",".txt",sep = ""))
```

#And to find the correlation between two kinship matrices
```{r}
cor(c(moa_K),c(back_K))
```