---
title: "VCAP Figures"
author: "Samantha Snodgrass"
date: "12/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Necessary Packages
```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)
```
Create a function to format the data files
```{r}
#assign(objname, read_tsv(p))
format_Hperm<-function(df){
  names(df)<- names(df) %>% str_replace_all(.,"_1","_SD")
  names(df)[13]<-"Trait"
  df<-df %>% select(-contains("Component"))
  return(df)
}
```

##MOA + Background 100 permutation test with all phenotypes
###Read in all files
```{r}
remove(list = ls(pattern = "^H_perm"))#so there isn't any confusion between test runs

for(i in 0:99){
  p<-str_c("1SNPperPeak_Test_Results/perm_",i,".h_estimates.txt",sep="")
  objname<-str_c("H_perm",i)
  assign(objname, read_tsv(p, col_types = cols()))
  assign(objname,format_Hperm(get(objname)))
}
```
###Merge all the individual permutation dfs into 1 main dataframe
```{r}
df_names<-ls(pattern = "^H_perm")
H_permALL<-H_perm0
for(i in 2:length(df_names)){
  H_permALL<-bind_rows(H_permALL, get(df_names[i]))
}

lookup<- c(Her_K1 = "Her_K1...1", Her_K2 = "Her_K2...2", Her_K3 = "Her_K3...3",
           Her_Top = "Her_Top...4", Her_All = "Her_All...5", SDHer_K1 = "Her_K1...6", SDHer_K2="Her_K2...7", SDHer_K3="Her_K3...8", SDHer_Top="Her_Top...9", SDHer_All="Her_All...10")
for(i in 0:99){
  objname<-str_c("H_perm",i)
  assign(objname,rename(get(objname), all_of(lookup)))
}
H_permALL<-rename(H_permALL, all_of(lookup))
```
###Calculate Va and plot
Va == Heritability_K[1 || 2 || 3]/Heritability_ALL
https://github.com/AndiKur4/MaizePal
```{r}

H_perm_ALL<-mutate(H_permALL, 
                  Va_MOA = Her_K1 / Her_All,
                  Va_BKGD = Her_K2 / Her_All,
                  Va_Rest = Her_K3 / Her_All)

melted_HpermALL<-H_perm_ALL %>% select(c(contains("Va"),Trait)) %>% melt(id.vars="Trait", value.name="percent_Va", variable.name="K_Component")
```

###Let's group phenotypes by "category"
```{r}
tassel_architecture<-c("tassel_length_BLUP;Brown_2011","spike_length_BLUP;Brown_2011","branch_zone_BLUP;Brown_2011","branch_number_transformed_BLUP;Brown_2011","tassprimbranchno_BLUP;Hung_2012_1","tasslength_BLUP;Hung_2012_1")
ear_architecture<-c("cob_length_BLUP;Brown_2011","cob_diameter_BLUP;Brown_2011","ear_row_number_transformed_BLUP;Brown_2011","cobdiam_BLUP;Hung_2012_1","coblength_BLUP;Hung_2012_1","earrowno_BLUP;Hung_2012_1","kernelnoperrow_BLUP;Hung_2012_1","earmass_BLUP;Hung_2012_1","cobmass_BLUP;Hung_2012_1","totalkernelweight_BLUP;Hung_2012_1","weight20kernels_BLUP;Hung_2012_1","totalkernelno_BLUP;Hung_2012_1","starch_kernel_BLUP;Cook_2012","protein_kernel_BLUP;Cook_2012","oil_kernel_BLUP;Cook_2012")
stalk_strength<-c("WI_rind_penetrometer_resistance_BLUE;Peiffer_2013","WI_rind_penetrometer_resistance_BLUP;Peiffer_2013","ALL_rind_penetrometer_resistance_BLUE;Peiffer_2013","ALL_rind_penetrometer_resistance_BLUP;Peiffer_2013","MO_rind_penetrometer_resistance_BLUE;Peiffer_2013","MO_rind_penetrometer_resistance_BLUP;Peiffer_2013","NY_rind_penetrometer_resistance_BLUE;Peiffer_2013","NY_rind_penetrometer_resistance_BLUP;Peiffer_2013")
disease<-c("southern_leaf_blight_BLUP;Bian_2014_Kump_2011","lesion_severity_BLUP;Olukolu_2014","sqrt_diseased_leaf_area1_BLUP;Poland_2011","sqrt_diseased_leaf_area2_BLUP;Poland_2011","sqrt_diseased_leaf_area3_BLUP;Poland_2011","diseased_leaf_area1_BLUP;Poland_2011","diseased_leaf_area2_BLUP;Poland_2011","diseased_leaf_area3_BLUP;Poland_2011","northern_leaf_blight_index_BLUP;Poland_2011")
plant_architecture<-c("height_ratio_BLUP;Olukolu_2014","stalk_width_ratio_BLUP;Olukolu_2014","leaf_length_BLUP;Tian_2011","leaf_width_BLUP;Tian_2011","upper_leaf_angle_BLUP;Tian_2011","last_leaf_with_epicuticular_wax;Foerster_2015","ALL_ear_height_BLUE;Peiffer_2013","ALL_ear_height_BLUP;Peiffer_2013","plantheight_BLUP;Hung_2012_1","earheight_BLUP;Hung_2012_1","leaflength_BLUP;Hung_2012_1","leafwidth_BLUP;Hung_2012_1","upperleafangle_BLUP;Hung_2012_1","nodenumberbelowear_BLUP;Hung_2012_1","nodenumberaboveear_BLUP;Hung_2012_1","numbraceroots_BLUP;Hung_2012_1","plant_height_BLUP;Peiffer_2014","ear_height_BLUP;Peiffer_2014","plant_height_minus_ear_height_BLUP;Peiffer_2014","ear_height_div_plant_height_BLUP;Peiffer_2014","plant_height_div_DTA_BLUP;Peiffer_2014","leaf_angle_boxcox_transformed_BLUP;Tian_2011")
flowering_time<-c("DTA_ratio_BLUP;Olukolu_2014","ALL_DTA_BLUE;Peiffer_2013","ALL_DTA_BLUP;Peiffer_2013","DTS_BLUP;Hung_2012_1","DTA_BLUP;Hung_2012_1","asi_BLUP;Hung_2012_1","DTA_BLUP;Buckler_2009","DTS_BLUP;Buckler_2009","DTA_BLUP;Wallace_2014","ASI_BLUP;Buckler_2009","GDD_DTS_BLUP;Peiffer_2014","GDD_DTA_BLUP;Peiffer_2014","GDD_ASI_BLUP;Peiffer_2014","DTS_BLUP;Peiffer_2014","DTA_BLUP;Peiffer_2014","ASI_BLUP;Peiffer_2014","GDD_DTA_long_BLUP;Hung_2012","GDD_DTA_short_BLUP;Hung_2012","GDD_DTA_photo_resp_BLUP;Hung_2012","GDD_DTS_long_BLUP;Hung_2012","GDD_DTS_short_BLUP;Hung_2012","GDD_DTS_photo_resp_BLUP;Hung_2012","GDD_DTA_NC06_BLUP;Hung_2012","GDD_DTS_NC06_BLUP;Hung_2012","GDD_DTA_MO06_BLUP;Hung_2012","GDD_DTS_MO06_BLUP;Hung_2012","GDD_DTA_NY06_BLUP;Hung_2012","GDD_DTS_NY06_BLUP;Hung_2012","GDD_DTA_IL06_BLUP;Hung_2012","GDD_DTS_IL06_BLUP;Hung_2012","GDD_DTA_FL06_BLUP;Hung_2012","GDD_DTS_FL06_BLUP;Hung_2012","GDD_DTA_PR06_BLUP;Hung_2012","GDD_DTS_PR06_BLUP;Hung_2012","GDD_DTA_NC07_BLUP;Hung_2012","GDD_DTS_NC07_BLUP;Hung_2012","GDD_DTA_MO07_BLUP;Hung_2012","GDD_DTS_MO07_BLUP;Hung_2012","GDD_DTA_NY07_BLUP;Hung_2012","GDD_DTS_NY07_BLUP;Hung_2012","GDD_DTA_IL07_BLUP;Hung_2012","GDD_DTS_IL07_BLUP;Hung_2012","GDD_DTA_FL07_BLUP;Hung_2012","GDD_DTS_FL07_BLUP;Hung_2012")
misc<-c("glsblup;Benson_2015","flecking_ls_mean;Olukolu_2016","flecking_poisson_transformation_BLUP;Olukolu_2016")
vitamin_E<-c("alphaT_transformed_BLUE;Diepenbrock_2017","deltaT_transformed_BLUE;Diepenbrock_2017","gammaT_transformed_BLUE;Diepenbrock_2017","TotalT_transformed_BLUE;Diepenbrock_2017","alphaT3_transformed_BLUE;Diepenbrock_2017","deltaT3_transformed_BLUE;Diepenbrock_2017","gammaT3_transformed_BLUE;Diepenbrock_2017","TotalT3_transformed_BLUE;Diepenbrock_2017","TotalT_plus_T3_transformed_BLUE;Diepenbrock_2017","plastochromanol8_transformed_BLUE;Diepenbrock_2017","alphaT_div_gammaT_transformed_BLUE;Diepenbrock_2017","alphaT3_div_gammaT3_transformed_BLUE;Diepenbrock_2017","gammaT_div_gammaTPlusalphaT_transformed_BLUE;Diepenbrock_2017","gammaT3_div_gammaT3PlusalphaT3_transformed_BLUE;Diepenbrock_2017","deltaT_div_gammaTPlusalphaT_transformed_BLUE;Diepenbrock_2017","deltaT_div_alphaT_transformed_BLUE;Diepenbrock_2017","deltaT_div_gammaT_transformed_BLUE;Diepenbrock_2017","deltaT3_div_gammaT3PlusalphaT3_transformed_BLUE;Diepenbrock_2017","deltaT3_div_alphaT3_transformed_BLUE;Diepenbrock_2017","deltaT3_div_gammaT3_transformed_BLUE;Diepenbrock_2017","TotalT_div_totalT3_transformed_BLUE;Diepenbrock_2017")
metabolites<-c("chlorophylla_BLUP;Wallace_2014","chlorophyllb_BLUP;Wallace_2014","malate_BLUP;Wallace_2014","fumarate_BLUP;Wallace_2014","fumarate2_BLUP;Wallace_2014","glutamate_BLUP;Wallace_2014","aminoacids_BLUP;Wallace_2014","protein_BLUP;Wallace_2014","nitrate_BLUP;Wallace_2014","starch_BLUP;Wallace_2014","sucrose_BLUP;Wallace_2014","glucose_BLUP;Wallace_2014","fructose_BLUP;Wallace_2014","principal_component_metabolies1;Wallace_2014","principal_component_metabolies2;Wallace_2014")
```
Splitting the plot by category
```{r}
melted_stats_HpermALL<-H_perm_ALL %>% select(c(contains("Va"),Trait)) %>% melt(id.vars="Trait", value.name="percent_Va", variable.name="K_Component") %>% group_by(Trait,K_Component) %>% summarize(Mean = mean(percent_Va), Median = median(percent_Va), std_dev=sd(percent_Va)) 
#can also use the above command to get the statistics for Va% for each component across all traits and by traits

melted_HpermALL<-mutate(melted_HpermALL, Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"))
melted_stats_HpermALL<-mutate(melted_stats_HpermALL, Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"))
```

####For MM poster
```{r}
H_perm_ALL<-mutate(H_perm_ALL, Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"),
                     trait_colors = case_when(Trait_Category == "ear_architecture" ~ "chocolate3",
                                              Trait_Category == "disease" ~ "darkgreen",
                                              Trait_Category == "stalk_strength" ~ "deepskyblue",
                                              Trait_Category == "plant_architecture" ~ "firebrick",
                                              Trait_Category == "flowering_time" ~ "goldenrod",
                                              Trait_Category == "misc" ~ "grey61",
                                              Trait_Category == "vitamin_E" ~ "navyblue",
                                              Trait_Category == "metabolites" ~ "purple",
                                              Trait_Category == "tassel_architecture" ~ "violetred1"))
H_perm_ALL<-arrange(H_perm_ALL, Va_MOA)

#MOA_AF<-read_tsv("Full_EG80WW_MOA_AF.distribution")
#BKGD_AF<-read_tsv("Full_EG80WW_BKGD_AF.distribution")

#ggplot()+
#  geom_bar(stat = "identity", data = BKGD_AF, aes(x=AFbin, y=AF_bin_count))+
#  geom_bar(stat = "identity", data = MOA_AF, fill = "red", aes(x=AFbin, y=AF_bin_count))+
#  xlab("Allele Frequency Bin (0.10 interval)")+ylab("SNPs per Bin (count)")
#ggsave("MM_poster_Combined_AFspectrum.png", device = "png", dpi = 300)
#ggplot()+
#geom_bar(stat = "identity", data = MOA_AF, fill = "red", aes(x=AFbin, y=AF_bin_count))+xlab("")+ylab("")
#ggsave("MM_poster_MOAonly_AFspectrum.png", device = "png", dpi = 300)
```

### ZEALUTION FIGURES ###
```{r}
library(ggridges)
#Uses MEAN for points
ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_MOA))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_MOA,fill="1_MOA"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2")+
  geom_point(data = melted_stats_HpermALL, aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_blank())+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("1SNPperPeak_ridgeplotVaALL_MEAN.png",device = "png")
##trying to get the traits labeled with colors
#need to create a vector that has the trait colors in order of VaMOA (needs to be just 1 length(It's not currently working because it's just doing all the permutations!!Not a unique list!))
Vatrait_label<-filter(melted_stats_HpermALL, K_Component == "Va_MOA") %>% arrange(., Mean) %>% mutate(., trait_colors = case_when(Trait_Category == "ear_architecture" ~ "chocolate3",
                                              Trait_Category == "disease" ~ "darkgreen",
                                              Trait_Category == "stalk_strength" ~ "deepskyblue",
                                              Trait_Category == "plant_architecture" ~ "firebrick",
                                              Trait_Category == "flowering_time" ~ "goldenrod",
                                              Trait_Category == "misc" ~ "grey61",
                                              Trait_Category == "vitamin_E" ~ "navyblue",
                                              Trait_Category == "metabolites" ~ "purple",
                                              Trait_Category == "tassel_architecture" ~ "violetred1")) %>% ungroup() %>% select(c(Trait_Category,trait_colors)) 
ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_MOA))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_MOA,fill="1_MOA"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2",labels = c("MOA", "Background", "Rest"))+
  geom_point(data = melted_stats_HpermALL, aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4), labels = c("MOA Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_text(color = Vatrait_label$trait_colors))+
  scale_y_discrete(label=Vatrait_label$Trait_Category)+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("1SNPperPeak_RidgeplotVaALL_MEAN_traitlabeled.png",device = "png")
###Uses MEDIAN for points
ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_MOA))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_MOA,fill="1_MOA"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2")+
  geom_point(data = melted_stats_HpermALL, aes(y=Trait, x=Median, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_blank())+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("1SNPperPeak_ridgeplotVaALL_MEDIAN.png",device = "png")

#Remaking with h2 instead of Va%
melted_HpermALL.Va<-melted_HpermALL
melted_HpermALL<-H_perm_ALL %>% select(c(Her_K1, Her_K2,Her_K3,Her_All,Trait)) %>% melt(id.vars="Trait", value.name="h2", variable.name="K_Component")

melted_stats_HpermALL<-H_perm_ALL %>% select(c(Her_K1, Her_K2,Her_K3,Her_All,Trait)) %>% 
  melt(id.vars="Trait", value.name="h2", variable.name="K_Component") %>%
  group_by(Trait,K_Component) %>% 
  summarize(Mean = mean(h2), Median = median(h2), std_dev=sd(h2)) 
h2trait_label<-filter(melted_stats_HpermALL, K_Component == "Her_K1") %>% arrange(., Mean) %>% mutate(., Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"),
                                  trait_colors = case_when(Trait_Category == "ear_architecture" ~ "chocolate3",
                                              Trait_Category == "disease" ~ "darkgreen",
                                              Trait_Category == "stalk_strength" ~ "deepskyblue",
                                              Trait_Category == "plant_architecture" ~ "firebrick",
                                              Trait_Category == "flowering_time" ~ "goldenrod",
                                              Trait_Category == "misc" ~ "grey61",
                                              Trait_Category == "vitamin_E" ~ "navyblue",
                                              Trait_Category == "metabolites" ~ "purple",
                                              Trait_Category == "tassel_architecture" ~ "violetred1")) %>% ungroup() %>% select(c(Trait_Category,trait_colors)) 
ggplot(H_perm_ALL, aes(y=reorder(Trait,Her_K1))) + 
  stat_density_ridges(aes(x=Her_K3,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Her_K2,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Her_K1,fill="1_MOA"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("h2") +
  scale_fill_brewer(palette = "Dark2",labels = c("MOA", "Background", "Rest"))+
  geom_point(data = filter(melted_stats_HpermALL, K_Component != "Her_All"), aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4), labels =c("MOA Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_text(color = Vatrait_label$trait_colors))+
  scale_y_discrete(label=Vatrait_label$Trait_Category)+
  ggtitle(str_c("Estimated h2 of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("1SNPperPeak_Ridgeplot.h2.all.traitlabeled.png", device = "png")

filter(melted_HpermALL, Trait %in% c("plantheight_BLUP;Hung_2012_1","plant_height_BLUP;Peiffer_2014","earheight_BLUP;Hung_2012_1","ear_height_BLUP;Peiffer_2014", "ALL_ear_height_BLUP;Peiffer_2013","ALL_ear_height_BLUE;Peiffer_2013") & K_Component != "Her_ALL") %>%
  group_by(Trait) %>% 
  ggplot(aes(x=Trait)) + geom_boxplot(aes(y=h2, fill = K_Component, color = K_Component))+
  coord_flip()+theme(axis.text.y = element_text(size = 12))+
  xlab("")+ylab("h2")+
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Dark2")+
  ggtitle("Height Trait Subset")+
  theme_minimal()#+theme(panel.grid.major.y = element_blank())
ggsave("1SNPperPeak_boxplot.heightonly.png", device = "png", height=7.45,width = 12.36,units = "in")
```
Looking at Vitamin E h2
```{r}
VitE.h2<-filter(melted_HpermALL, Trait %in% vitamin_E) %>% mutate(Component = case_when(K_Component == "Her_K1" ~ "MOA", K_Component == "Her_K2" ~ "BKGD", K_Component == "Her_K3" ~ "REST", K_Component == "Her_ALL" ~ "Total h2"))
VitE.h2$Component<- factor(VitE.h2$Component, levels = c("MOA","BKGD","REST","Total h2"))

plastochromanol<-c("plastochromanol8")
tocopherol<-c("alphaT","deltaT","gammaT","TotalT","alphaT_div_gammaT","gammaT_div_gammaTPlusalphaT","deltaT_div_gammaTPlusalphaT","deltaT_div_alphaT","deltaT_div_gammaT")
tocotrienol<-c("alphaT3","deltaT3","gammaT3","TotalT3","alphaT3_div_gammaT3","gammaT3_div_gammaT3PlusalphaT3","deltaT3_div_gammaT3PlusalphaT3","deltaT3_div_alphaT3","deltaT3_div_gammaT3")
tocochromanol<-c("TotalT_plus_T3","TotalT_div_totalT3")

VitE.h2<-mutate(VitE.h2, bettertraitname = str_remove_all(Trait, "_transformed_BLUE;Diepenbrock_2017"),
       pathway = case_when(bettertraitname %in% plastochromanol ~ "plastochromanol",
                           bettertraitname %in% tocopherol ~ "tocopherol",
                           bettertraitname %in% tocotrienol ~ "tocotrienol",
                           bettertraitname %in% tocochromanol ~ "tocochromanol"))

ggplot(VitE.h2,aes(x = Trait, y= h2))+
  geom_bar(data = filter(VitE.h2, Component == "Total h2") %>%
              group_by(Trait) %>% summarize(Mean = mean(h2)), 
           aes(x=Trait, y=Mean),stat = "identity")+
  geom_boxplot(data = filter(VitE.h2, Component != "Total h2"), aes(fill = Component, color = Component))+
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Dark2")+
  coord_flip()+
  theme_minimal()+
  ggtitle("Vitamin E estimated h2")

VitE.h2.stats<-VitE.h2 %>% group_by(Trait,Component,pathway) %>% summarize(Mean = mean(h2), Median = median(h2))
#dcast(VitE.h2.stats, Trait ~ Component, value.var = "Mean")  %>% ggplot(aes(x=reorder(Trait, `Total h2`)))+
#  geom_bar(stat = "identity", aes(y=`Total h2`), color = "black")+
#  geom_point(aes(y=REST), color = "purple") +
#  geom_point(aes(y = BKGD), color = "orange") + 
#  geom_point(aes(y = MOA), color = "green")+
#  coord_flip()+
#  theme_minimal()+
#  ggtitle("Vitamin E mean h2 estimates")
VitE.h2.stats$Trait<-VitE.h2.stats$Trait %>% str_replace(pattern = "_transformed_BLUE;Diepenbrock_2017",replacement = "") %>% str_replace_all(pattern = "_",replacement = " ") %>% str_replace_all(pattern = "Plus", replacement = " Plus ")
VitE.h2.stats %>% filter(Component != "Total h2") %>% 
  ggplot(aes(fill = Component, y=Mean, x = reorder(Trait, Mean)))+
  geom_bar(position="stack", stat="identity")+
  ggtitle("Mean Vitamin E estimates")+
  theme_minimal()+
  scale_x_discrete(labels = function(Trait) str_wrap(Trait, width = 10))+
  #coord_flip()+
  facet_wrap(facets = vars(pathway), scales = "free_x")+
  theme(strip.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size=6))+
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Dark2")+ xlab("Traits")+ ylab("h2")
#ggsave("VitaminE.h2byComponent.barchart.png", device = "png", dpi = 300, width = 9, height = 6)

filter(VitE.h2, K_Component != "Her_All") %>%
ggplot(aes(y=bettertraitname)) + 
  stat_density_ridges(aes(x=h2,fill=K_Component),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("h2") +
  scale_fill_brewer(palette = "Dark2",labels = c("MOA", "Background", "Rest"))+
  #geom_point(data = filter(VitE.h2.stats, Component != "Total h2"), aes(y=Trait, x=Mean, shape = Component)) + 
  #scale_shape_manual(values=c(16, 6, 4), labels =c("MOA Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme( axis.text.y = element_blank(), panel.border = element_rect(color = "black", fill = NA, size = 1), strip.background = element_rect(color="black", size=1.5, linetype="solid"), strip.text.y = element_text(face = "bold"))+
  facet_grid(facets = vars(pathway), scales = "free_y")+
  ggtitle(str_c("Estimated h2 of Vitamin E traits"))+
  theme(legend.title = element_blank())
ggsave("1SNPperPeak_VitaminE.h2byComponent.ridgeplot.png", device = "png")
```
Creating Supplemental table with h2 estimates for 143 traits
```{r}
#melted_stats_HpermALL<-H_perm_ALL %>% select(c(Her_K1, Her_K2,Her_K3,Her_ALL,Trait)) %>% melt(id.vars="Trait", value.name="h2", variable.name="K_Component") %>%group_by(Trait,K_Component) %>% summarize(Mean = mean(h2), Median = median(h2), std_dev=sd(h2))
#cast long data into wide frame
h2stats.mean<-dcast(melted_stats_HpermALL, Trait ~  K_Component, value.var = "Mean")
h2stats.median<-dcast(melted_stats_HpermALL, Trait ~  K_Component, value.var = "Median")
h2stats.stddev<-dcast(melted_stats_HpermALL, Trait ~  K_Component, value.var = "std_dev")
#change column names ahead of joining
colnames(h2stats.mean)[2:5]<-colnames(h2stats.mean)[2:5] %>% str_c(.,".mean")
colnames(h2stats.median)[2:5]<-colnames(h2stats.median)[2:5] %>% str_c(.,".median")
colnames(h2stats.stddev)[2:5]<-colnames(h2stats.stddev)[2:5] %>% str_c(.,".StdDev")
#join datatables
h2.table<-full_join(x=h2stats.mean, y=h2stats.median,by="Trait") %>% full_join(x=., y=h2stats.stddev, by="Trait")
#Replace Numbers with Component Names
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_K1", replacement = "MOA")
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_K2", replacement = "BKGD")
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_K3", replacement = "REST")
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_All", replacement = "TOTAL")
#Order
h2.table<-h2.table %>% select(c(Trait, contains("MOA"),contains("BKGD"),contains("REST"),contains("TOTAL")))
#Save table
write_csv(h2.table,"SupplementalTable.1SNPperPeak.h2estimates.alltraitsbycomponent.csv")
```
Is heritability estimates for MOA correlated with overall heritability?
```{r}
#cor(x=h2.table$MOA.mean, y=h2.table$TOTAL.mean)
#correlation between MOA and TOTAL
cor(x=filter(melted_HpermALL, K_Component == "Her_K1")%>%select(h2), 
    y=filter(melted_HpermALL, K_Component == "Her_All")%>%select(h2))
#correlation between BKGD and TOTAL
cor(x=filter(melted_HpermALL, K_Component == "Her_K2")%>%select(h2), 
    y=filter(melted_HpermALL, K_Component == "Her_All")%>%select(h2))
#correlation between Rest and TOTAL
cor(x=filter(melted_HpermALL, K_Component == "Her_K3")%>%select(h2), 
    y=filter(melted_HpermALL, K_Component == "Her_All")%>%select(h2))
```
MOA 1 SNP per peak is correlated to overall heritability of the trait at 0.4475, while the background is correlated at 0.3413 and the rest at 0.0965.  

##Testing for significant differences in Va by K_component and trait using the 1 SNP per peak MOA results
Testing it out as a general
```{r}
model.1 <- aov(percent_Va ~ K_Component + Trait, data= melted_HpermALL.Va)
summary(model.1)
TukeyHSD(model.1,which = "K_Component")
```
So, this means that for %Va, the general MOA explain on average 13% more than the background and 39% more than the Rest. 

What about on average across all traits, what is Va% for bQTL vs. background vs. rest?

```{r}
melted_HpermALL.Va %>% filter(K_Component == "Va_MOA") %>% summarize(MOA_mean = mean(percent_Va, na.rm = TRUE))

melted_HpermALL.Va %>% filter(K_Component == "Va_BKGD") %>% summarize(BKGD_mean = mean(percent_Va, na.rm = TRUE))

melted_HpermALL.Va %>% filter(K_Component == "Va_Rest") %>% summarize(Rest_mean = mean(percent_Va, na.rm = TRUE))
```

Testing it by trait
```{r}
#Loop to test for differences in TPM by Pangene and MOA SNP
trait.list<-H_perm_ALL$Trait %>% unique() 

aov.p<-c()
trait<-c()
f.stat<-c()
trait<-c()
con.trait<-c()
con.name<-c()
con.p<-c()
con.diff<-c()

for(i in 1:length(trait.list)){
  t<-filter(melted_HpermALL.Va, Trait == trait.list[i])
  agep<-aov(percent_Va ~ K_Component + 0, data = t)
  con<-TukeyHSD(agep, which = "K_Component")
  aov.p<-c(aov.p, summary(agep)[[1]][[5]][[1]])
  f.stat<-c(f.stat, summary(agep)[[1]][[4]][[1]])
  trait<-c(trait,trait.list[i])
  con.trait<-c(con.trait,trait.list[i],trait.list[i],trait.list[i] )
  con.name<-c(con.name, con$K_Component %>% dimnames()%>% .[[1]])
  con.p<-c(con.p, con$K_Component[,4]) %>% unname()
  con.diff<-c(con.diff, con$K_Component[,1]) %>% unname()
}
ANOVA.Results<-tibble(P_val = aov.p, F_Stat =f.stat,Trait = trait)
Contrast.Results<-tibble(Trait = con.trait, Contrast_Name = con.name, P_val = con.p ,Contrast_Difference = con.diff)
```
Check for significance?
```{r}
nrow(ANOVA.Results)
0.05/143 #Bonferroni cut off
#Which of the tests meet the Bonferroni cut off?
Bonferroni.aov<-filter(ANOVA.Results, P_val <=  0.0003496503)

###Calculate Benjamini-Hochberg values for test values
ANOVA.Results<-arrange(ANOVA.Results, P_val) #arrange p-values smallest to largest
#Find the largest integer k so that p(k) <= k*alpha/m
BH_05<-c()
for(k in 1:nrow(ANOVA.Results)){
  temp<-(k*0.05)/nrow(ANOVA.Results)
  BH_05<-c(BH_05,temp)
}
ANOVA.Results<-add_column(ANOVA.Results,BH_05)
ANOVA.Results
BH05.aov<-filter(ANOVA.Results, P_val <= ANOVA.Results$BH_05)

```
Which traits were significantly different? All of them, no matter the corrective cutoff.

```{r}
melted_stats_Va<-H_perm_ALL %>% select(c(contains("Va"),Trait)) %>% melt(id.vars=c("Trait"), value.name="percent_Va", variable.name="K_Component") %>% group_by(Trait,K_Component) %>% summarize(Mean = mean(percent_Va), Median = median(percent_Va), std_dev=sd(percent_Va)) 

filter(Contrast.Results, P_val >= 0.05) #Contrasts for traits that weren't significant
filter(Contrast.Results, P_val >= 0.05 & Contrast_Name %in% c("Va_Rest-Va_bQTL", "Va_BKGD-Va_MOA"))
notSigTraits<-filter(Contrast.Results, P_val >= 0.05 & Contrast_Name %in% c("Va_Rest-Va_bQTL", "Va_BKGD-Va_MOA")) %>% select(Trait)
filter(melted_stats_Va, Trait %in% notSigTraits$Trait)
filter(Contrast.Results, Contrast_Name == "Va_BKGD-Va_MOA" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the MOA estimate was larger than the Background estimated Va and significant
filter(Contrast.Results, Contrast_Name == "Va_Rest-Va_MOA" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the MOA estimate was larger than the rest estimated Va and significant
filter(Contrast.Results, Contrast_Name == "Va_Rest-Va_BKGD" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the background estimate was larger than the rest estimated Va and significant
```
1. There were 20 contrasts that were insignificant 
2. 11 of which were contrasts involving the MOA component
3. Which corresponds to 11 traits where MOA was not different from either the background or rest component
4. There were 84 traits where the MOA component accounted for more %Va than the background
5. There were 124 traits where the MOA component accounted for more %Va than the rest
6. There were 116 traits where the background component accounted for more %Va than the rest

How many traits had a mean percent Va for bQTL that was 0.5 or higher?
```{r}
melted_stats_Va %>% filter(K_Component == "Va_MOA" & Mean >= 0.5) %>% nrow()
80/143
```
There were 80 traits where the mean percent Va for general MOA was 50% or higher, which is ~56% of traits. 

Remaking stats with h2 instead of Va%
```{r}
model.h2 <- aov(h2 ~ K_Component + Trait, data= filter(melted_HpermALL, K_Component != "Her_All"))
summary(model.h2)
TukeyHSD(model.h2,which = "K_Component")
```
Interesting that here trait is significant in the anova whereas for the %Va it is not. 

On average, MOA had 0.06 more h2 than background and 0.199 more than rest. The background had 0.14 more h2 than the rest. 
```{r}
#Loop to test for differences by trait

aov.p<-c()
trait<-c()
f.stat<-c()
trait<-c()
con.trait<-c()
con.name<-c()
con.p<-c()
con.diff<-c()

for(i in 1:length(trait.list)){
  t<-filter(melted_HpermALL, Trait == trait.list[i] & K_Component != "Her_All")
  agep<-aov(h2 ~ K_Component + 0, data = t)
  con<-TukeyHSD(agep, which = "K_Component")
  aov.p<-c(aov.p, summary(agep)[[1]][[5]][[1]])
  f.stat<-c(f.stat, summary(agep)[[1]][[4]][[1]])
  trait<-c(trait,trait.list[i])
  con.trait<-c(con.trait,trait.list[i],trait.list[i],trait.list[i] )
  con.name<-c(con.name, con$K_Component %>% dimnames()%>% .[[1]])
  con.p<-c(con.p, con$K_Component[,4]) %>% unname()
  con.diff<-c(con.diff, con$K_Component[,1]) %>% unname()
}
ANOVA.Results.h2<-tibble(P_val = aov.p, F_Stat =f.stat,Trait = trait)
Contrast.Results.h2<-tibble(Trait = con.trait, Contrast_Name = con.name, P_val = con.p ,Contrast_Difference = con.diff)

nrow(ANOVA.Results.h2)
0.05/143 #Bonferroni cut off
#Which of the tests meet the Bonferroni cut off?
Bonferroni.h2.aov<-filter(ANOVA.Results.h2, P_val <=  0.0003496503) 

###Calculate Benjamini-Hochberg values for test values
ANOVA.Results.h2<-arrange(ANOVA.Results.h2, P_val) #arrange p-values smallest to largest
#Find the largest integer k so that p(k) <= k*alpha/m
BH_05.h2<-c()
for(k in 1:nrow(ANOVA.Results.h2)){
  temp<-(k*0.05)/nrow(ANOVA.Results.h2)
  BH_05.h2<-c(BH_05.h2,temp)
}
ANOVA.Results.h2<-add_column(ANOVA.Results.h2,BH_05.h2)
ANOVA.Results.h2
BH05.h2.aov<-filter(ANOVA.Results.h2, P_val <= ANOVA.Results.h2$BH_05.h2)
```
How many traits pass significance threshold? All of them are significant no matter the corrective threshold.

```{r}
filter(Contrast.Results.h2, P_val >= 0.05) #Contrasts for traits that weren't significant
filter(Contrast.Results.h2, P_val >= 0.05 & Contrast_Name %in% c("Her_K3-Her_K1", "Her_K2-Her_K1")) #K1 = MOA, K2 = BKGD, K3= Rest
notSigTraits<-filter(Contrast.Results.h2, P_val >= 0.05 & Contrast_Name %in% c("Her_K3-Her_K1", "Her_K2-Her_K1")) %>% select(Trait)
filter(melted_stats_HpermALL, Trait %in% notSigTraits$Trait)
filter(Contrast.Results.h2, Contrast_Name == "Her_K2-Her_K1" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the MOA estimate was larger than the Background 
filter(Contrast.Results.h2, Contrast_Name == "Her_K3-Her_K1" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the MOA estimate was larger than the rest 
# How often was background estimates higher than rest
filter(Contrast.Results.h2, Contrast_Name == "Her_K3-Her_K2" & P_val <= 0.05 & Contrast_Difference < 0)
```
1. There were 22 contrasts that weren't significant
2. Of those, 14 involved the MOA component
3. These represent 13 traits (TotalT_plus_T3 from Diepenbrock 2017 had 2 insignificant contrasts)
4. 82 traits where the MOA accounted for more h2 than the background
5. 124 traits where the MOA accounted for more h2 than the rest
6. 116 traits where the background accounted for more h2 than the rest

##bQTL Strict VCAP Results
###Read in all files
```{r}
remove(list = ls(pattern = "^H_perm"))#so there isn't any confusion between test runs

for(i in 0:99){
  p<-str_c("bQTL_Strict_Oct2023/Test_Results/perm_",i,".h_estimates.txt",sep="")
  objname<-str_c("H_perm",i)
  assign(objname, read_tsv(p, col_types = cols()))
  assign(objname,format_Hperm(get(objname)))
}

lookup<- c(Her_K1 = "Her_K1...2", Her_K2 = "Her_K2...3", Her_K3 = "Her_K3...4",
           Her_Top = "Her_Top...5", Her_All = "Her_All...6", SDHer_K1 = "Her_K1...8", SDHer_K2="Her_K2...9", SDHer_K3="Her_K3...10", SDHer_Top="Her_Top...11", SDHer_All="Her_All...12")
for(i in 0:99){
  objname<-str_c("H_perm",i)
  assign(objname,rename(get(objname), all_of(lookup)))
}
```
###Merge all the individual permutation dfs into 1 main dataframe
```{r}
df_names<-ls(pattern = "^H_perm")
H_permALL<-H_perm0
for(i in 2:length(df_names)){
  H_permALL<-bind_rows(H_permALL, get(df_names[i]))
}
```
###Calculate Va and plot
Va == Heritability_K[1 || 2 || 3]/Heritability_ALL
https://github.com/AndiKur4/MaizePal
```{r}

H_perm_ALL<-mutate(H_permALL, 
                  Va_bQTL = Her_K1 / Her_All,
                  Va_BKGD = Her_K2 / Her_All,
                  Va_Rest = Her_K3 / Her_All)

melted_HpermALL<-H_perm_ALL %>% select(c(contains("Va"),Trait)) %>% melt(id.vars="Trait", value.name="percent_Va", variable.name="K_Component")
```

Splitting the plot by category
```{r}
melted_stats_HpermALL<-H_perm_ALL %>% select(c(contains("Va"),Trait)) %>% melt(id.vars="Trait", value.name="percent_Va", variable.name="K_Component") %>% group_by(Trait,K_Component) %>% summarize(Mean = mean(percent_Va), Median = median(percent_Va), std_dev=sd(percent_Va)) 
#can also use the above command to get the statistics for Va% for each component across all traits and by traits

melted_HpermALL<-mutate(melted_HpermALL, Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"))
melted_stats_HpermALL<-mutate(melted_stats_HpermALL, Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"))
```
####For MM poster
```{r}
H_perm_ALL<-mutate(H_perm_ALL, Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"),
                     trait_colors = case_when(Trait_Category == "ear_architecture" ~ "chocolate3",
                                              Trait_Category == "disease" ~ "darkgreen",
                                              Trait_Category == "stalk_strength" ~ "deepskyblue",
                                              Trait_Category == "plant_architecture" ~ "firebrick",
                                              Trait_Category == "flowering_time" ~ "goldenrod",
                                              Trait_Category == "misc" ~ "grey61",
                                              Trait_Category == "vitamin_E" ~ "navyblue",
                                              Trait_Category == "metabolites" ~ "purple",
                                              Trait_Category == "tassel_architecture" ~ "violetred1"))
H_perm_ALL<-arrange(H_perm_ALL, Va_bQTL)

bQTL_AF<-read_tsv("bQTL_Strict_Oct2023/bQTL_AF_dist.strict.txt", col_names = c("Chr.snp","Start.snp","Stop.snp","AF","distance","Chr.clump","Start.clump","Stop.clump","Clump.ID"))
BKGD_AF<-read_tsv("bQTL_Strict_Oct2023/nonbQTL_AF_dist.strict.txt", col_names = c("Chr.snp","Start.snp","Stop.snp","AF","distance"))

bQTL_AF<-mutate(bQTL_AF, AFbin = case_when(AF <= 0.10 ~ "AF.0-10",
                                      AF > 0.10 & AF <= 0.20 ~ "AF.10-20",
                                      AF > 0.20 & AF <= 0.30 ~ "AF.20-30",
                                      AF > 0.30 & AF <= 0.40 ~ "AF.30-40",
                                      AF > 0.40 & AF <= 0.50 ~ "AF.40-50",
                                      AF > 0.50 & AF <= 0.60 ~ "AF.50-60",
                                      AF > 0.60 & AF <= 0.70 ~ "AF.60-70",
                                      AF > 0.70 & AF <= 0.80 ~ "AF.70-80",
                                      AF > 0.80 & AF <= 0.90 ~ "AF.80-90",
                                      AF > 0.90 & AF <= 1 ~ "AF.90-100"),
                    Distbin = case_when(distance <= 500 & distance >= -500 ~ "dist.0-500",
                                        distance <= 1000 & distance > 500 ~ "dist.500-1000",
                                        distance >= -1000 & distance < -500 ~ "dist.500-1000",
                                        distance <= 5000 & distance > 1000 ~ "dist.1000-5000",
                                        distance >= -5000 & distance < -1000 ~ "dist.1000-5000",
                                        distance > 5000 | distance < -5000 ~ "dist.5000-max"))
BKGD_AF<-mutate(BKGD_AF, AFbin = case_when(AF <= 0.10 ~ "AF.0-10",
                                      AF > 0.10 & AF <= 0.20 ~ "AF.10-20",
                                      AF > 0.20 & AF <= 0.30 ~ "AF.20-30",
                                      AF > 0.30 & AF <= 0.40 ~ "AF.30-40",
                                      AF > 0.40 & AF <= 0.50 ~ "AF.40-50",
                                      AF > 0.50 & AF <= 0.60 ~ "AF.50-60",
                                      AF > 0.60 & AF <= 0.70 ~ "AF.60-70",
                                      AF > 0.70 & AF <= 0.80 ~ "AF.70-80",
                                      AF > 0.80 & AF <= 0.90 ~ "AF.80-90",
                                      AF > 0.90 & AF <= 1 ~ "AF.90-100"),
                    Distbin = case_when(distance <= 500 & distance >= -500 ~ "dist.0-500",
                                        distance <= 1000 & distance > 500 ~ "dist.500-1000",
                                        distance >= -1000 & distance < -500 ~ "dist.500-1000",
                                        distance <= 5000 & distance > 1000 ~ "dist.1000-5000",
                                        distance >= -5000 & distance < -1000 ~ "dist.1000-5000",
                                        distance > 5000 | distance < -5000 ~ "dist.5000-max"))

ggplot()+
  geom_bar(data = BKGD_AF, aes(x=AFbin), stat="count")+
  xlab("Allele Frequency Bin (0.10 interval)")+ylab("SNPs per Bin (count)")+
  theme_minimal()+ggtitle("AF of bgSNP pool")
ggsave("bQTLStrict-Oct2023-bgSNP_AFspectrum.png", device = "png", dpi = 300)
ggplot()+
geom_bar(data = bQTL_AF, fill = "red", aes(x=AFbin), stat="count")+
  xlab("Allele Frequency Bin (0.10 interval)")+ylab("SNPs per Bin (count)")+
  theme_minimal()+ggtitle("AF of bQTL SNP pool")
ggsave("bQTLStrict-Oct2023-bQTL_AFspectrum.png", device = "png", dpi = 300)

BKGD_AF$Distbin <- factor(BKGD_AF$Distbin, levels = c("dist.0-500","dist.500-1000","dist.1000-5000","dist.5000-max"))
bQTL_AF$Distbin <- factor(bQTL_AF$Distbin, levels = c("dist.0-500","dist.500-1000","dist.1000-5000","dist.5000-max"))
ggplot()+
  geom_bar(data=BKGD_AF, aes(x=Distbin), stat = "count")+
  xlab("Distance to nearest gene (bp)")+ylab("SNPs per Bin (count)")+
  theme_minimal()+ggtitle("Gene distance of bgSNP pool")
ggsave("bQTLStrict-Oct2023-bgSNP_distancebins.png", device = "png", dpi = 300)
ggplot()+
  geom_bar(data=bQTL_AF, aes(x=Distbin),fill = "red", stat = "count")+
  xlab("Distance to nearest gene (bp)")+ylab("SNPs per Bin (count)")+
  theme_minimal()+ggtitle("Gene distance of bQTL pool")
ggsave("bQTLStrict-Oct2023-bQTL_distancebins.png", device = "png", dpi = 300)

```

### ZEALUTION FIGURES ###
```{r}
library(ggridges)
#Uses MEAN for points
ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_bQTL))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_bQTL,fill="1_bQTL"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2")+
  geom_point(data = melted_stats_HpermALL, aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_blank())+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("bQTLstrict_ridgeplotVaALL_MEAN.png",device = "png")
##trying to get the traits labeled with colors
#need to create a vector that has the trait colors in order of VaMOA (needs to be just 1 length(It's not currently working because it's just doing all the permutations!!Not a unique list!))
Vatrait_label<-filter(melted_stats_HpermALL, K_Component == "Va_bQTL") %>% arrange(., Mean) %>% mutate(., trait_colors = case_when(Trait_Category == "ear_architecture" ~ "chocolate3",
                                              Trait_Category == "disease" ~ "darkgreen",
                                              Trait_Category == "stalk_strength" ~ "deepskyblue",
                                              Trait_Category == "plant_architecture" ~ "firebrick",
                                              Trait_Category == "flowering_time" ~ "goldenrod",
                                              Trait_Category == "misc" ~ "grey61",
                                              Trait_Category == "vitamin_E" ~ "navyblue",
                                              Trait_Category == "metabolites" ~ "purple",
                                              Trait_Category == "tassel_architecture" ~ "violetred1")) %>% ungroup() %>% select(c(Trait_Category,trait_colors)) 
ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_bQTL))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_bQTL,fill="1_bQTL"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2",labels = c("bQTL", "Background", "Rest"))+
  geom_point(data = melted_stats_HpermALL, aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4), labels = c("bQTL Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_text(color = Vatrait_label$trait_colors))+
  scale_y_discrete(label=Vatrait_label$Trait_Category)+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("bQTLstrict_RidgeplotVaALL_MEAN_traitlabeled.png",device = "png")
library(svglite)
ggsave("bQTLstrict_RidgeplotVaALL_MEAN_traitlabeled.svg",device = "svg")

###Uses MEDIAN for points
ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_bQTL))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_bQTL,fill="1_bQTL"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2")+
  geom_point(data = filter(melted_stats_HpermALL, K_Component != "Her_All"), aes(y=Trait, x=Median, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_blank())+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("bQTLstrict_ridgeplotVaALL_MEDIAN.png",device = "png")

#Remaking with h2 instead of Va%
melted_HpermALL.Va<-melted_HpermALL
melted_HpermALL<-H_perm_ALL %>% select(c(Her_K1, Her_K2,Her_K3,Her_All,Trait)) %>% melt(id.vars="Trait", value.name="h2", variable.name="K_Component")

melted_stats_HpermALL<-H_perm_ALL %>% select(c(Her_K1, Her_K2,Her_K3,Her_All,Trait)) %>% 
  melt(id.vars="Trait", value.name="h2", variable.name="K_Component") %>%
  group_by(Trait,K_Component) %>% 
  summarize(Mean = mean(h2), Median = median(h2), std_dev=sd(h2)) 
h2trait_label<-filter(melted_stats_HpermALL, K_Component == "Her_K1") %>% arrange(., Mean) %>% mutate(., Trait_Category = case_when(Trait %in% tassel_architecture ~ "tassel_architecture",
                                  Trait %in% ear_architecture ~ "ear_architecture",
                                  Trait %in% disease ~ "disease",
                                  Trait %in% stalk_strength ~ "stalk_strength",
                                  Trait %in% plant_architecture ~ "plant_architecture",
                                  Trait %in% flowering_time ~ "flowering_time",
                                  Trait %in% misc ~ "misc",
                                  Trait %in% vitamin_E ~ "vitamin_E",
                                  Trait %in% metabolites ~ "metabolites"),
                                  trait_colors = case_when(Trait_Category == "ear_architecture" ~ "chocolate3",
                                              Trait_Category == "disease" ~ "darkgreen",
                                              Trait_Category == "stalk_strength" ~ "deepskyblue",
                                              Trait_Category == "plant_architecture" ~ "firebrick",
                                              Trait_Category == "flowering_time" ~ "goldenrod",
                                              Trait_Category == "misc" ~ "grey61",
                                              Trait_Category == "vitamin_E" ~ "navyblue",
                                              Trait_Category == "metabolites" ~ "purple",
                                              Trait_Category == "tassel_architecture" ~ "violetred1")) %>% ungroup() %>% select(c(Trait_Category,trait_colors)) 
ggplot(H_perm_ALL, aes(y=reorder(Trait,Her_K1))) + 
  stat_density_ridges(aes(x=Her_K3,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Her_K2,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Her_K1,fill="1_bQTL"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("h2") +
  scale_fill_brewer(palette = "Dark2",labels = c("bQTL", "Background", "Rest"))+
  geom_point(data = filter(melted_stats_HpermALL, K_Component != "Her_All"), aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4), labels =c("bQTL Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(), axis.text.y = element_text(color = Vatrait_label$trait_colors))+
  scale_y_discrete(label=Vatrait_label$Trait_Category)+
  ggtitle(str_c("Estimated h2 of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank())
ggsave("bQTLstrict_Ridgeplot.h2.all.traitlabeled.png", device = "png")
ggsave("bQTLstrict_Ridgeplot.h2.all.traitlabeled.svg", device = "svg")

filter(melted_HpermALL, Trait %in% c("plantheight_BLUP;Hung_2012_1","plant_height_BLUP;Peiffer_2014","earheight_BLUP;Hung_2012_1","ear_height_BLUP;Peiffer_2014", "ALL_ear_height_BLUP;Peiffer_2013","ALL_ear_height_BLUE;Peiffer_2013") & K_Component != "Her_All") %>%
  group_by(Trait) %>% 
  ggplot(aes(x=Trait)) + geom_boxplot(aes(y=h2, fill = K_Component, color = K_Component))+
  coord_flip()+theme(axis.text.y = element_text(size = 12))+
  xlab("")+ylab("h2")+
  scale_fill_brewer(palette = "Set2", labels=c("bQTL","BKGD","Rest"))+
  scale_color_brewer(palette = "Dark2",labels=c("bQTL","BKGD","Rest"))+
  ggtitle("Height Trait Subset")+
  theme_minimal()+theme(panel.grid.major.y = element_blank())
ggsave("bQTLstrict_boxplot.heightonly.png", device = "png", height=7.45,width = 12.36,units = "in")
```
Looking at Vitamin E h2
```{r}
VitE.h2<-filter(melted_HpermALL, Trait %in% vitamin_E) %>% mutate(Component = case_when(K_Component == "Her_K1" ~ "bQTL", K_Component == "Her_K2" ~ "BKGD", K_Component == "Her_K3" ~ "REST", K_Component == "Her_ALL" ~ "Total h2"))
VitE.h2$Component<- factor(VitE.h2$Component, levels = c("bQTL","BKGD","REST","Total h2"))

plastochromanol<-c("plastochromanol8")
tocopherol<-c("alphaT","deltaT","gammaT","TotalT","alphaT_div_gammaT","gammaT_div_gammaTPlusalphaT","deltaT_div_gammaTPlusalphaT","deltaT_div_alphaT","deltaT_div_gammaT")
tocotrienol<-c("alphaT3","deltaT3","gammaT3","TotalT3","alphaT3_div_gammaT3","gammaT3_div_gammaT3PlusalphaT3","deltaT3_div_gammaT3PlusalphaT3","deltaT3_div_alphaT3","deltaT3_div_gammaT3")
tocochromanol<-c("TotalT_plus_T3","TotalT_div_totalT3")

VitE.h2<-mutate(VitE.h2, bettertraitname = str_remove_all(Trait, "_transformed_BLUE;Diepenbrock_2017"),
       pathway = case_when(bettertraitname %in% plastochromanol ~ "plastochromanol",
                           bettertraitname %in% tocopherol ~ "tocopherol",
                           bettertraitname %in% tocotrienol ~ "tocotrienol",
                           bettertraitname %in% tocochromanol ~ "tocochromanol"))

ggplot(VitE.h2,aes(x = Trait, y= h2))+
  geom_bar(data = filter(VitE.h2, Component == "Total h2") %>%
              group_by(Trait) %>% summarize(Mean = mean(h2)), 
           aes(x=Trait, y=Mean),stat = "identity")+
  geom_boxplot(data = filter(VitE.h2, Component != "Total h2"), aes(fill = Component, color = Component))+
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Dark2")+
  coord_flip()+
  theme_minimal()+
  ggtitle("Vitamin E estimated h2")

VitE.h2.stats<-VitE.h2 %>% group_by(Trait,Component,pathway) %>% summarize(Mean = mean(h2), Median = median(h2))
#dcast(VitE.h2.stats, Trait ~ Component, value.var = "Mean")  %>% ggplot(aes(x=reorder(Trait, `Total h2`)))+
#  geom_bar(stat = "identity", aes(y=`Total h2`), color = "black")+
#  geom_point(aes(y=REST), color = "purple") +
#  geom_point(aes(y = BKGD), color = "orange") + 
#  geom_point(aes(y = MOA), color = "green")+
#  coord_flip()+
#  theme_minimal()+
#  ggtitle("Vitamin E mean h2 estimates")
VitE.h2.stats$Trait<-VitE.h2.stats$Trait %>% str_replace(pattern = "_transformed_BLUE;Diepenbrock_2017",replacement = "") %>% str_replace_all(pattern = "_",replacement = " ") %>% str_replace_all(pattern = "Plus", replacement = " Plus ")
VitE.h2.stats %>% filter(Component != "Total h2") %>% 
  ggplot(aes(fill = Component, y=Mean, x = reorder(Trait, Mean)))+
  geom_bar(position="stack", stat="identity")+
  ggtitle("Mean Vitamin E estimates")+
  theme_minimal()+
  scale_x_discrete(labels = function(Trait) str_wrap(Trait, width = 10))+
  #coord_flip()+
  facet_wrap(facets = vars(pathway), scales = "free_x")+
  theme(strip.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size=6))+
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Dark2")+ xlab("Traits")+ ylab("h2")
ggsave("bQTLstrict.VitaminE.h2byComponent.barchart.png", device = "png", dpi = 300, width = 9, height = 6)

filter(VitE.h2, K_Component != "Her_All") %>%
ggplot(aes(y=bettertraitname)) + 
  stat_density_ridges(aes(x=h2,fill=K_Component),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("h2") +
  scale_fill_brewer(palette = "Dark2",labels = c("bQTL", "Background", "Rest"))+
  #geom_point(data = filter(VitE.h2.stats, Component != "Total h2"), aes(y=Trait, x=Mean, shape = Component)) + 
  #scale_shape_manual(values=c(16, 6, 4), labels =c("MOA Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme( axis.text.y = element_blank(), panel.border = element_rect(color = "black", fill = NA, size = 1), strip.background = element_rect(color="black", size=1.5, linetype="solid"), strip.text.y = element_text(face = "bold"))+
  facet_grid(facets = vars(pathway), scales = "free_y")+
  ggtitle(str_c("Estimated h2 of Vitamin E traits"))+
  theme(legend.title = element_blank())
ggsave("bQTLstrict_VitaminE.h2byComponent.ridgeplot.png", device = "png", height = 7.5, width = 7.5)
```
##Making the publication quality figures
Ridgeplots
```{r}
#%Va ridgeplot for the bQTL
melted_stats_Va<-H_perm_ALL %>% select(c(contains("Va"),Trait)) %>% melt(id.vars=c("Trait"), value.name="percent_Va", variable.name="K_Component") %>% group_by(Trait,K_Component) %>% summarize(Mean = mean(percent_Va), Median = median(percent_Va), std_dev=sd(percent_Va)) 

ggplot(H_perm_ALL, aes(y=reorder(Trait,Va_bQTL))) + 
  stat_density_ridges(aes(x=Va_Rest,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_BKGD,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Va_bQTL,fill="1_bQTL"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("Va %") +
  scale_fill_brewer(palette = "Dark2",labels = c("bQTL", "Background", "Rest"))+
  geom_point(data = melted_stats_Va, aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4), labels = c("bQTL Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme(panel.grid = element_blank(), axis.text.y = element_text(color = Vatrait_label$trait_colors, face = "bold"))+
  scale_y_discrete(label=rep("",143))+
  scale_x_continuous(expand = c(0,0))+
  ggtitle(str_c("Estimated Va of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank(),
        legend.position = "none")
ggsave("pub_VaRidgeplot.svg",device = "svg",width = 4.32, height=5.52)
ggsave("pub_VaRidgeplot.png",device = "png")

#h2 ridgeplot for the bQTL
ggplot(H_perm_ALL, aes(y=reorder(Trait,Her_K1))) + 
  stat_density_ridges(aes(x=Her_K3,fill="3_Rest"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Her_K2,fill="2_Background"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  stat_density_ridges(aes(x=Her_K1,fill="1_bQTL"),alpha=0.75,size=0,from=0,to=1,bandwidth = 0.01)+
  ylab("Trait")+ xlab("h2") +
  scale_fill_brewer(palette = "Dark2",labels = c("bQTL", "Background", "Rest"))+
  geom_point(data = filter(melted_stats_HpermALL, K_Component != "Her_All"), aes(y=Trait, x=Mean, shape = K_Component)) + 
  scale_shape_manual(values=c(16, 6, 4), labels = c("bQTL Mean", "Background Mean", "Rest Mean"))+
  theme_minimal()+
  theme(panel.grid = element_blank(), axis.text.y = element_text(color = Vatrait_label$trait_colors, face = "bold"))+
  scale_y_discrete(label=rep("",143))+
  scale_x_continuous(expand = c(0,0))+
  ggtitle(str_c("Estimated h2 of All Traits with 100 Permutations"))+
  theme(legend.title = element_blank(),
        legend.position = "none")
ggsave("pub_h2Ridgeplot.svg",device = "svg",width = 4.32, height=5.52)
ggsave("pub_h2Ridgeplot.png",device = "png")
```
Boxplot with a subset of traits
```{r}
#Find the traits to include
# 1 leaf trait
plant_architecture[grepl("*leaf*", plant_architecture)] # put it in line 958
# 1 flowering time trait
# 1 disease trait
#probably want to use either "northern_leaf_blight_index_BLUP;Poland_2011" or "southern_leaf_blight_BLUP;Bian_2014_Kump_2011"
# 1 low performing trait (Vitamin E)
traitsubset.Va<-filter(melted_HpermALL.Va, Trait %in% c("plantheight_BLUP;Hung_2012_1","plant_height_BLUP;Peiffer_2014","leaflength_BLUP;Hung_2012_1","ALL_DTA_BLUP;Peiffer_2013", "DTS_BLUP;Peiffer_2014", "southern_leaf_blight_BLUP;Bian_2014_Kump_2011", "deltaT_transformed_BLUE;Diepenbrock_2017") & K_Component != "Her_All")
traitsubset.Va$Trait <- factor(traitsubset.Va$Trait, levels = c("plantheight_BLUP;Hung_2012_1","plant_height_BLUP;Peiffer_2014","leaflength_BLUP;Hung_2012_1","ALL_DTA_BLUP;Peiffer_2013", "DTS_BLUP;Peiffer_2014", "southern_leaf_blight_BLUP;Bian_2014_Kump_2011", "deltaT_transformed_BLUE;Diepenbrock_2017"))
traitsubset.Va<-mutate(traitsubset.Va, plotting_names = case_when(Trait == "plantheight_BLUP;Hung_2012_1" ~ "Plant Height, Hung 2012",
                                                                  Trait == "plant_height_BLUP;Peiffer_2014" ~ "Plant Height, Peiffer 2014",
                                                                  Trait == "leaflength_BLUP;Hung_2012_1" ~ "Leaf length, Hung 2012", 
                                                                  Trait == "ALL_DTA_BLUP;Peiffer_2013" ~ "DTA, Peiffer 2013", 
                                                                  Trait == "DTS_BLUP;Peiffer_2014" ~ "DTS, Peiffer 2014",
                                                                  Trait == "southern_leaf_blight_BLUP;Bian_2014_Kump_2011" ~ "S. leaf blight, Bian 2014, Kump 2011",
                                                                  Trait == "deltaT_transformed_BLUE;Diepenbrock_2017" ~ "delta T, Diepenbrock 2017" ))
traitsubset.Va$plotting_names <- factor(traitsubset.Va$plotting_names, levels = c("Plant Height, Hung 2012","Plant Height, Peiffer 2014","Leaf length, Hung 2012","DTA, Peiffer 2013","DTS, Peiffer 2014", "S. leaf blight, Bian 2014, Kump 2011","delta T, Diepenbrock 2017"))
traitsubset.Va %>%
#filter(melted_HpermALL, Trait %in% vitamin_E & K_Component != "Her_All") %>%
  group_by(plotting_names) %>% 
  ggplot(aes(x=plotting_names)) + geom_boxplot(aes(y=percent_Va, color = K_Component))+
  theme_minimal()+
  theme(axis.text.y = element_text(size = 12), 
        axis.text.x = element_text(angle = 90),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")+
  scale_x_discrete(labels = function(plotting_names) str_wrap(plotting_names, width = 15))+
  xlab(" ")+ylab("Va%")+
  #scale_fill_brewer(palette = "Set2", labels=c("bQTL","BKGD","Rest"))+
  scale_color_brewer(palette = "Set2",labels=c("bQTL","BKGD","Rest"))+
  ggtitle("Subset of traits using bQTL SNPs")
ggsave("pub_VaTraitSubsetBoxplot.svg",device = "svg", height = 4, width = 5)
ggsave("pub_VaTraitSubsetBoxplot.png",device = "png", height = 4, width = 5)
```

Creating Supplemental table with h2 estimates for 143 traits
```{r}
#melted_stats_HpermALL<-H_perm_ALL %>% select(c(Her_K1, Her_K2,Her_K3,Her_ALL,Trait)) %>% melt(id.vars="Trait", value.name="h2", variable.name="K_Component") %>%group_by(Trait,K_Component) %>% summarize(Mean = mean(h2), Median = median(h2), std_dev=sd(h2))
#cast long data into wide frame
h2stats.mean<-dcast(melted_stats_HpermALL, Trait ~  K_Component, value.var = "Mean")
h2stats.median<-dcast(melted_stats_HpermALL, Trait ~  K_Component, value.var = "Median")
h2stats.stddev<-dcast(melted_stats_HpermALL, Trait ~  K_Component, value.var = "std_dev")
#change column names ahead of joining
colnames(h2stats.mean)[2:5]<-colnames(h2stats.mean)[2:5] %>% str_c(.,".mean")
colnames(h2stats.median)[2:5]<-colnames(h2stats.median)[2:5] %>% str_c(.,".median")
colnames(h2stats.stddev)[2:5]<-colnames(h2stats.stddev)[2:5] %>% str_c(.,".StdDev")
#join datatables
h2.table<-full_join(x=h2stats.mean, y=h2stats.median,by="Trait") %>% full_join(x=., y=h2stats.stddev, by="Trait")
#Replace Numbers with Component Names
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_K1", replacement = "bQTL")
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_K2", replacement = "BKGD")
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_K3", replacement = "REST")
colnames(h2.table)<-colnames(h2.table) %>% str_replace_all(pattern = "Her_All", replacement = "TOTAL")
#Order
h2.table<-h2.table %>% select(c(Trait, contains("bQTL"),contains("BKGD"),contains("REST"),contains("TOTAL")))
#Save table
write_csv(h2.table,"SupplementalTable.bQTLstrict.h2estimates.alltraitsbycomponent.csv")
```
Is heritability estimates for MOA correlated with overall heritability?
```{r}
#correlation between bQTL and TOTAL
cor(x=filter(melted_HpermALL, K_Component == "Her_K1")%>%select(h2), 
    y=filter(melted_HpermALL, K_Component == "Her_All")%>%select(h2), method = "pearson")
#correlation between BKGD and TOTAL
cor(x=filter(melted_HpermALL, K_Component == "Her_K2")%>%select(h2), 
    y=filter(melted_HpermALL, K_Component == "Her_All")%>%select(h2), method = "pearson")
#correlation between Rest and TOTAL
cor(x=filter(melted_HpermALL, K_Component == "Her_K3")%>%select(h2), 
    y=filter(melted_HpermALL, K_Component == "Her_All")%>%select(h2), method = "pearson")
```

##Testing for significant differences in Va by K_component and trait using the strict bQTL results
Testing it out as a general
```{r}
model.1 <- aov(percent_Va ~ K_Component + Trait, data= melted_HpermALL.Va)
summary(model.1)
TukeyHSD(model.1,which = "K_Component")
```
So, this means that for %Va, the bQTL explain on average 29.9% more than the background and 49.7% more than the Rest. 

What about on average across all traits, what is Va% for bQTL vs. background vs. rest?

```{r}
melted_HpermALL.Va %>% filter(K_Component == "Va_bQTL") %>% summarize(bQTL_mean = mean(percent_Va, na.rm = TRUE))

melted_HpermALL.Va %>% filter(K_Component == "Va_BKGD") %>% summarize(BKGD_mean = mean(percent_Va, na.rm = TRUE))

melted_HpermALL.Va %>% filter(K_Component == "Va_Rest") %>% summarize(Rest_mean = mean(percent_Va, na.rm = TRUE))
```

Testing it by trait
```{r}
#Loop to test for differences in TPM by Pangene and MOA SNP
trait.list<-H_perm_ALL$Trait %>% unique() 

aov.p<-c()
trait<-c()
f.stat<-c()
trait<-c()
con.trait<-c()
con.name<-c()
con.p<-c()
con.diff<-c()

#t<-filter(melted_HpermALL, Trait == trait.list[1])
#agep<-aov(percent_Va ~ K_Component + 0, data = t)
#con<-TukeyHSD(agep, which = "K_Component")
#aov.p<-c(aov.p, summary(agep)[[1]][[5]][[1]])
#f.stat<-c(f.stat, summary(agep)[[1]][[4]][[1]])
#trait<-c(trait,trait.list[1])
#con.trait<-c(con.trait,trait.list[1],trait.list[1],trait.list[1] )
#con.name<-c(con.name, con$K_Component %>% dimnames()%>% .[[1]])
#con.p<-c(con.p, con$K_Component[,4]) %>% unname()
#con.diff<-c(con.diff, con$K_Component[,1]) %>% unname()


for(i in 1:length(trait.list)){
  t<-filter(melted_HpermALL.Va, Trait == trait.list[i])
  agep<-aov(percent_Va ~ K_Component + 0, data = t)
  con<-TukeyHSD(agep, which = "K_Component")
  aov.p<-c(aov.p, summary(agep)[[1]][[5]][[1]])
  f.stat<-c(f.stat, summary(agep)[[1]][[4]][[1]])
  trait<-c(trait,trait.list[i])
  con.trait<-c(con.trait,trait.list[i],trait.list[i],trait.list[i] )
  con.name<-c(con.name, con$K_Component %>% dimnames()%>% .[[1]])
  con.p<-c(con.p, con$K_Component[,4]) %>% unname()
  con.diff<-c(con.diff, con$K_Component[,1]) %>% unname()
}
ANOVA.Results<-tibble(P_val = aov.p, F_Stat =f.stat,Trait = trait)
Contrast.Results<-tibble(Trait = con.trait, Contrast_Name = con.name, P_val = con.p ,Contrast_Difference = con.diff)
```
Check for significance?
```{r}
nrow(ANOVA.Results)
0.05/143 #Bonferroni cut off
#Which of the tests meet the Bonferroni cut off?
Bonferroni.aov<-filter(ANOVA.Results, P_val <=  0.0003496503)

###Calculate Benjamini-Hochberg values for test values
ANOVA.Results<-arrange(ANOVA.Results, P_val) #arrange p-values smallest to largest
#Find the largest integer k so that p(k) <= k*alpha/m
BH_05<-c()
for(k in 1:nrow(ANOVA.Results)){
  temp<-(k*0.05)/nrow(ANOVA.Results)
  BH_05<-c(BH_05,temp)
}
ANOVA.Results<-add_column(ANOVA.Results,BH_05)
ANOVA.Results
BH05.aov<-filter(ANOVA.Results, P_val <= ANOVA.Results$BH_05)

```
All the traits were significantly different when tested for percent_VA ~ K_component + 0 no matter which correction was used.
```{r}
filter(Contrast.Results, P_val >= 0.05) #Contrasts for traits that weren't significant
filter(Contrast.Results, P_val >= 0.05 & Contrast_Name %in% c("Va_Rest-Va_bQTL", "Va_BKGD-Va_bQTL"))
notSigTraits<-filter(Contrast.Results, P_val >= 0.05 & Contrast_Name %in% c("Va_Rest-Va_bQTL", "Va_BKGD-Va_bQTL")) %>% select(Trait)
filter(melted_stats_Va, Trait %in% notSigTraits$Trait)
filter(Contrast.Results, Contrast_Name == "Va_BKGD-Va_bQTL" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the bQTL estimate was larger than the Background estimated Va and significant
filter(Contrast.Results, Contrast_Name == "Va_Rest-Va_bQTL" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the bQTL estimate was larger than the rest estimated Va and significant
filter(Contrast.Results, Contrast_Name == "Va_Rest-Va_BKGD" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the background estimate was larger than the rest estimated Va and significant
```
1. There were 15 contrasts that were not significantly different
2. Of those 15, 13 were not significant differences between the background and the bQTL components or the rest and the bQTL components. (The other two were insignificant differences between the background and rest component)
3. 39/3 = 13 = the number of traits that weren't significant. This pulls out the other stats for them too. 
4. There were 106 traits where the bQTL estimated Va was larger than the background and significantly different
5. There were 123 traits where the bQTL estimated Va was larger than the rest and significantly different.
6. There were 121 traits where the background estimated Va was larger than the rest and significantly different. 

How many traits had a mean percent Va for bQTL that was 0.5 or higher?
```{r}
melted_stats_Va %>% filter(K_Component == "Va_bQTL" & Mean >= 0.5) %>% nrow()
101/143
```
There were 101 traits where the mean percent Va for bQTL was 50% or higher, which is ~70% of traits. 

Remaking stats with h2 instead of Va%
```{r}
model.h2 <- aov(h2 ~ K_Component + Trait, data= filter(melted_HpermALL, K_Component != "Her_All"))
summary(model.h2)
TukeyHSD(model.h2,which = "K_Component")
```

```{r}
#Loop to test for differences by trait

aov.p<-c()
trait<-c()
f.stat<-c()
trait<-c()
con.trait<-c()
con.name<-c()
con.p<-c()
con.diff<-c()

for(i in 1:length(trait.list)){
  t<-filter(melted_HpermALL, Trait == trait.list[i] & K_Component != "Her_All")
  agep<-aov(h2 ~ K_Component + 0, data = t)
  con<-TukeyHSD(agep, which = "K_Component")
  aov.p<-c(aov.p, summary(agep)[[1]][[5]][[1]])
  f.stat<-c(f.stat, summary(agep)[[1]][[4]][[1]])
  trait<-c(trait,trait.list[i])
  con.trait<-c(con.trait,trait.list[i],trait.list[i],trait.list[i] )
  con.name<-c(con.name, con$K_Component %>% dimnames()%>% .[[1]])
  con.p<-c(con.p, con$K_Component[,4]) %>% unname()
  con.diff<-c(con.diff, con$K_Component[,1]) %>% unname()
}
ANOVA.Results.h2<-tibble(P_val = aov.p, F_Stat =f.stat,Trait = trait)
Contrast.Results.h2<-tibble(Trait = con.trait, Contrast_Name = con.name, P_val = con.p ,Contrast_Difference = con.diff)

nrow(ANOVA.Results.h2)
0.05/143 #Bonferroni cut off
#Which of the tests meet the Bonferroni cut off?
Bonferroni.h2.aov<-filter(ANOVA.Results.h2, P_val <=  0.0003496503) 

###Calculate Benjamini-Hochberg values for test values
ANOVA.Results.h2<-arrange(ANOVA.Results.h2, P_val) #arrange p-values smallest to largest
#Find the largest integer k so that p(k) <= k*alpha/m
BH_05.h2<-c()
for(k in 1:nrow(ANOVA.Results.h2)){
  temp<-(k*0.05)/nrow(ANOVA.Results.h2)
  BH_05.h2<-c(BH_05.h2,temp)
}
ANOVA.Results.h2<-add_column(ANOVA.Results.h2,BH_05.h2)
ANOVA.Results.h2
BH05.h2.aov<-filter(ANOVA.Results.h2, P_val <= ANOVA.Results.h2$BH_05.h2)
```
All traits are significantly different by component for h2 no matter the type of correction used. 

```{r}
filter(Contrast.Results.h2, P_val >= 0.05) #Contrasts for traits that weren't significant
filter(Contrast.Results.h2, P_val >= 0.05 & Contrast_Name %in% c("Her_K3-Her_K1", "Her_K2-Her_K1")) #K1 = MOA, K2 = BKGD, K3= Rest
notSigTraits<-filter(Contrast.Results.h2, P_val >= 0.05 & Contrast_Name %in% c("Her_K3-Her_K1", "Her_K2-Her_K1")) %>% select(Trait)
filter(melted_stats_HpermALL, Trait %in% notSigTraits$Trait)
filter(Contrast.Results.h2, Contrast_Name == "Her_K2-Her_K1" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the bQTL estimate was larger than the Background 
filter(Contrast.Results.h2, Contrast_Name == "Her_K3-Her_K1" & Contrast_Difference <= 0 & P_val <= 0.05) # traits where the bQTL estimate was larger than the rest 
# How often was background estimates higher than rest
filter(Contrast.Results.h2, Contrast_Name == "Her_K3-Her_K2" & P_val <= 0.05 & Contrast_Difference < 0)
```
1. There were 15 contrasts that weren't significant
2. Of those, 13 were contrasts between the bQTL component and either the background or rest.
3. Again, pulling out the stats of those 14 traits that weren't significant from #2
4. There were 106 traits where the bQTL accounted for significantly more h2 than background
5. There were 123 traits where the bQTL accounted for significantly more h2 than rest
6. There were 121 traits where the background accounted for significantly more h2 than rest

#### Simulation Test
Formatting the results for easier graphing
```{r}
Simulation_Results<-read_tsv("../JeffsTest/bQTL_strict_Oct2023/bQTL_Sim.h_estimates.txt")

Simulation_Results<-filter(Simulation_Results, Component == "Heritability")
colnames(Simulation_Results)[7]<-"Simulated_Trait"

Simulation_Results<-select(Simulation_Results, -Component) %>% melt(value.name = "Estimated_h2", id.vars = "Simulated_Trait", variable.name = "Her_Component") %>% filter(Her_Component != "Her_Top")

Simulation_Results<-mutate(Simulation_Results, Component = case_when(Her_Component == "Her_K1" ~ "bQTL", Her_Component == "Her_K2" ~ "bkgd", Her_Component == "Her_K3" ~ "rest", Her_Component == "Her_All" ~ "TOTAL"))

Simulation_Results<-mutate(Simulation_Results, True_h2 = case_when(Simulated_Trait %in% c(paste("Set1-",1:10,sep = "")) & Component == "bQTL" ~ 0.1,
                                               Simulated_Trait %in% c(paste("Set1-",1:10,sep = "")) & Component == "bkgd" ~ 0.5,
                                               Simulated_Trait %in% c(paste("Set1-",1:10,sep = "")) & Component == "rest" ~ 0.1,
                                               Simulated_Trait %in% c(paste("Set1-",1:10,sep = "")) & Component == "TOTAL" ~ 0.6,
                                               Simulated_Trait %in% c(paste("Set2-",1:10,sep = "")) & Component == "bQTL" ~ 0.5,
                                               Simulated_Trait %in% c(paste("Set2-",1:10,sep = "")) & Component == "bkgd" ~ 0.1,
                                               Simulated_Trait %in% c(paste("Set2-",1:10,sep = "")) & Component == "rest" ~ 0.1,
                                               Simulated_Trait %in% c(paste("Set2-",1:10,sep = "")) & Component == "TOTAL" ~ 0.6,
                                               Simulated_Trait %in% c(paste("Set3-",1:10,sep = "")) & Component == "bQTL" ~ 0.05,
                                               Simulated_Trait %in% c(paste("Set3-",1:10,sep = "")) & Component == "bkgd" ~ 0.05,
                                               Simulated_Trait %in% c(paste("Set3-",1:10,sep = "")) & Component == "rest" ~ 0.3,
                                               Simulated_Trait %in% c(paste("Set3-",1:10,sep = "")) & Component == "TOTAL" ~ 0.4,
                                               Simulated_Trait %in% c(paste("Set4-",1:10,sep = "")) & Component == "bQTL" ~ 0.9,
                                               Simulated_Trait %in% c(paste("Set4-",1:10,sep = "")) & Component == "bkgd" ~ 0.02,
                                               Simulated_Trait %in% c(paste("Set4-",1:10,sep = "")) & Component == "rest" ~ 0.02,
                                               Simulated_Trait %in% c(paste("Set4-",1:10,sep = "")) & Component == "TOTAL" ~ 0.94
                                               ))

Simulation_Results$Component <- factor(Simulation_Results$Component, levels = c("bQTL","bkgd","rest","TOTAL"))

```

Correlation between estimated h2 and true h2
```{r}
cor(x=Simulation_Results$True_h2, y=Simulation_Results$Estimated_h2)
cor(x=Simulation_Results$True_h2, y=Simulation_Results$Estimated_h2)^2 #Is this r^2?
```
Plot the results
```{r}
Simulation_Results %>% filter(Component %in% c("bQTL","bkgd","rest")) %>%
  mutate(Setofh2 = case_when(Simulated_Trait %in% paste0("Set1-",1:10) ~ "Set1",
                             Simulated_Trait %in% paste0("Set2-",1:10) ~ "Set2",
                             Simulated_Trait %in% paste0("Set3-",1:10) ~ "Set3",
                             Simulated_Trait %in% paste0("Set4-",1:10) ~ "Set4")) %>%
  ggplot(aes(x = Setofh2,fill=Component, y = Estimated_h2))+
    geom_boxplot(aes(color=Component))+
    geom_point(aes(y=True_h2), shape = 3, position=position_dodge(width = 0.75) )+
    scale_color_brewer(palette = "Dark2")+
    scale_fill_brewer(palette = "Set2")+
    theme_bw()+
    ylab("Estimated h2")+xlab("Sets of Simulated Phenotypes with Same h2 Truths")
ggsave("../JeffsTest/bQTL_strict_Oct2023/Simulatedh2Boxplot.png",device = "png") 
```
